---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- Add a project state badge
See https://github.com/BCDevExchange/Our-Project-Docs/blob/master/discussion/projectstates.md
If you have bcgovr installed and you use RStudio, click the 'Insert BCDevex Badge' Addin. -->[![img](https://img.shields.io/badge/Lifecycle-Stable-97ca00)](https://github.com/bcgov/repomountie/blob/master/doc/lifecycle-badges.md)

# bcdroughtstatistics

The primary output of this package are RMarkdown products that create tables and maps of drought relevant statistics for the province of British Columbia. These RMarkdown drought statistics packages are aligned by natural resource region, and have been developed with the input from staff from these regions who are tasked with managing and monitoring the progression of drought. Drought statistics are compiled to demonstrate the severity of hydrologic drought within a particular natural resource management region. Also included is a 'Low Flow Advisory' RMarkdown product that shows which streams across the entire province are low, and thus experiencing moderate to severe hydrologic drought impacts.

### Features
This package features three functions that render three variations on the drought statistics product. Two of the functions return drought statistics products tailored for specific regions within British Columbia; the third function returns the Low Flow Advisory product.


### Installation

You can install bcdroughtstatistics() from this Github repository using the remotes package though install.packages("remotes").

```{r install}
remotes::install_github("bcgov/bcdroughtstatistics")
library(bcdroughtstatistics)
```

### Usage

#### Example

An html version of the drought statistics package can be generated for a specific natural resource management region in BC 

```{r example}

library(bcdroughtstatistics)

# All of the NRM regions in BC
basins <- c("Cariboo Natural Resource Region",
            "Kootenay-Boundary Natural Resource Region",
            "Northeast Natural Resource Region",
            "Omineca Natural Resource Region",
            "Skeena Natural Resource Region",
            "South Coast Natural Resource Region",
            "Thompson-Okanagan Natural Resource Region",
            "West Coast Natural Resource Region")

## Get the Drought Statistics page for the West Coast without saving

drought_stats_wc(basin = "West Coast Natural Resource Region", save = FALSE)
```


### Project Status
In development.

### Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/bcgov/bcdroughtstatistics/issues/).

### How to Contribute

If you would like to contribute to the package, please see our 
[CONTRIBUTING](CONTRIBUTING.md) guidelines.

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

### License

```
Copyright 2021 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the &quot;License&quot;);
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an &quot;AS IS&quot; BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
```

---
*This project was created using the [bcgovr](https://github.com/bcgov/bcgovr) package.* 
